% A LaTeX template for MSc Thesis submissions to 
% Politecnico di Milano (PoliMi) - School of Industrial and Information Engineering
%
% S. Bonetti, A. Gruttadauria, G. Mescolini, A. Zingaro
% e-mail: template-tesi-ingind@polimi.it
%
% Last Revision: October 2021
%
% Copyright 2021 Politecnico di Milano, Italy. NC-BY

\documentclass{Configuration_Files/PoliMi3i_thesis}

%------------------------------------------------------------------------------
%	REQUIRED PACKAGES AND  CONFIGURATIONS
%------------------------------------------------------------------------------

% CONFIGURATIONS
\usepackage{parskip} % For paragraph layout
\usepackage{setspace} % For using single or double spacing
\usepackage{emptypage} % To insert empty pages
\usepackage{multicol} % To write in multiple columns (executive summary)
\setlength\columnsep{15pt} % Column separation in executive summary
\setlength\parindent{0pt} % Indentation

\raggedbottom  

% PACKAGES FOR TITLES
\usepackage{titlesec}
% \titlespacing{\section}{left spacing}{before spacing}{after spacing}
\titlespacing{\section}{0pt}{3.3ex}{2ex}
\titlespacing{\subsection}{0pt}{3.3ex}{1.65ex}
\titlespacing{\subsubsection}{0pt}{3.3ex}{1ex}
\usepackage{color}

% PACKAGES FOR LANGUAGE AND FONT
\usepackage[english]{babel} % The document is in English  
\usepackage[utf8]{inputenc} % UTF8 encoding
\usepackage[T1]{fontenc} % Font encoding
\usepackage[11pt]{moresize} % Big fonts

% PACKAGES FOR IMAGES
\usepackage{graphicx}
\usepackage{transparent} % Enables transparent images
\usepackage{eso-pic} % For the background picture on the title page
\usepackage{subfig} % Numbered and caption subfigures using \subfloat.
\usepackage{tikz} % A package for high-quality hand-made figures.
\usetikzlibrary{}
\graphicspath{{./Images/}} % Directory of the images
\usepackage{caption} % Coloured captions
\usepackage{xcolor} % Coloured captions
\usepackage{amsthm,thmtools,xcolor} % Coloured "Theorem"
\usepackage{float}

% STANDARD MATH PACKAGES
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage[overload]{empheq} % For braced-style systems of equations.
\usepackage{fix-cm} % To override original LaTeX restrictions on sizes

% PACKAGES FOR TABLES
\usepackage{tabularx}
\usepackage{longtable} % Tables that can span several pages
\usepackage{colortbl}

% PACKAGES FOR ALGORITHMS (PSEUDO-CODE)
\usepackage{algorithm}
\usepackage{algorithmic}

% PACKAGES FOR REFERENCES & BIBLIOGRAPHY
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref} % Adds clickable links at references
\usepackage{cleveref}
\usepackage[square, numbers, sort&compress]{natbib} % Square brackets, citing references with numbers, citations sorted by appearance in the text and compressed
%\bibliographystyle{unsrt} % You may use a different style adapted to your field
\bibliographystyle{ieeetr}

% OTHER PACKAGES
\usepackage{pdfpages} % To include a pdf file
\usepackage{afterpage}
\usepackage{lipsum} % DUMMY PACKAGE
\usepackage{fancyhdr} % For the headers
\fancyhf{}

% Input of configuration file. Do not change config.tex file unless you really know what you are doing. 
\input{Configuration_Files/config}


%----------------------------------------------------------------------------
%	NEW COMMANDS DEFINED
%----------------------------------------------------------------------------

% EXAMPLES OF NEW COMMANDS
\newcommand{\bea}{\begin{eqnarray}} % Shortcut for equation arrays
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\e}[1]{\times 10^{#1}}  % Powers of 10 notation
\newcommand{\mathbbm}[1]{\text{\usefont{U}{bbm}{m}{n}#1}} % From mathbbm.sty

\newcommand{\etal}{\textit{et al}., }
\newcommand{\ie}{\textit{i}.\textit{e}., }
\newcommand{\eg}{\textit{e}.\textit{g}.\ }
\newcommand{\wrt}{\textit{w}.\textit{r}.\textit{t}.\ }

%----------------------------------------------------------------------------
%	ADD YOUR PACKAGES (be careful of package interaction)
%----------------------------------------------------------------------------
\usepackage{easyReview}
\usepackage[utf8]{inputenc}
\usepackage[acronym, nomain, nonumberlist, nogroupskip]{glossaries}
\makeglossaries
\usepackage{mathtools}

%----------------------------------------------------------------------------
%	ADD YOUR DEFINITIONS AND COMMANDS (be careful of existing commands)
%----------------------------------------------------------------------------

%----------------------------------------------------------------------------
%	BEGIN OF YOUR DOCUMENT
%----------------------------------------------------------------------------

\begin{document}

\fancypagestyle{plain}{%
\fancyhf{} % Clear all header and footer fields
\fancyhead[RO,RE]{\thepage} %RO=right odd, RE=right even
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}
\input{acronyms}
%----------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------

\pagestyle{empty} % No page numbers
\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the preamble pages

\puttitle{
	title=Feasibility of Non-Line of Sight \newline Integrated Sensing and \newline Communication at mmWave, % Title of the thesis
	name=Paolo Tosi, % Author Name and Surname
	course=Telecommunications Engineering - Ingegneria delle Telecomunicazioni, % Study Programme (in Italian)
	ID  = 997038,  % Student ID number (numero di matricola)
	advisor= Prof. Maurizio Magarini, % Supervisor name
	coadvisor={Marcus Henninger, Dr. Silvio Mandelli}, % Co-Supervisor name, remove this line if there is none
	academicyear={2023-24},  % Academic Year
} % These info will be put into your Title page 

%----------------------------------------------------------------------------
%	PREAMBLE PAGES: ABSTRACT (inglese e italiano), EXECUTIVE SUMMARY
%----------------------------------------------------------------------------
\startpreamble
\setcounter{page}{1} % Set page counter to 1

% ABSTRACT IN ENGLISH
\chapter*{Abstract} 

One of the new key features behind the development of sixth-generation (6G) mobile networks is their ability to gather information about the surrounding environment. 
Orthogonal Frequency Division Multiplexing (OFDM) radio signals, which are transmitted by base stations to users and vice versa, not only carry data, but are also affected by the movement of the transmitter, obstacles, and illuminated targets.
The concept of OFDM radar aims to extract information about the distance and relative speeds of targets that reflect the signal.
The high bandwidths and large antenna apertures available in next-generation mobile networks will facilitate the usage of OFDM radar for gathering sensing information. 
The combination of these functions and the underlying existing communication capabilities is referred to as Integrated Sensing and Communication (ISAC).
Applications typically considered for ISAC include vehicle detection (cars or drones), human activity detection and industrial safety. 
The typical assumption is that the target is in \gls{los} from the sensing hardware. 
However, an area that has not been extensively explored is non-line of sight (NLOS) sensing, where the radio signal can, if at all, reach the target after multiple reflections.
The aim of this thesis work is to investigate the feasibility of ISAC in NLOS environments, addressing performance constraints imposed by communication requirements, algorithmic development and identification of novel application scenarios for the system, with a focus on intrusion detection in factory floors.
The proposed approaches are validated using real-world measurements obtained from Nokia's ISAC mmWave prototype, which was deployed in an industrial test facility in the ARENA 2036 in Stuttgart, Germany.
The prototype is based on a commercial system that operates at 27.4 GHz and complies with the 3rd Generation Partnership Project (3GPP) standards for fifth-generation (5G) waveforms in the mmWave frequency bands. 
% The system is extended by an additional radio unit called 'Sniffer' for receiving signals reflected by the target, and a 'Sensing Processing Unit (SPU)' which is responsible for performing the radar processing algorithms.
Validation measurements were performed focusing on human intrusion detection scenarios in an industrial environment.
The objective is to define an initial approach for NLOS radar sensing based on periodogram spectral analysis, which can be exploited for more advanced sensing functionalities.
\\
\\
\textbf{Keywords:} isac, 6G, ofdm radar, monostatic sensing, non-line-of-sight sensing % Keywords

% ABSTRACT IN ITALIAN
\chapter*{Abstract in lingua italiana}
Una delle nuove caratteristiche chiave alla base dello sviluppo delle reti mobili di sesta generazione (6G) sarà la loro capacità di raccogliere informazioni sull'ambiente circostante. 
I segnali radio a divisione di frequenza ortogonale (OFDM), che vengono trasmessi dalle stazioni base agli utenti e viceversa, non solo trasportano dati, ma sono anche influenzati dal movimento del trasmettitore, dagli ostacoli e dai soggetti illuminati dal segnale.
Il concetto di radar OFDM mira a estrarre informazioni sulla distanza e sulla velocità re-\linebreak~lativa dei bersagli analizzando il segnale riflesso.
Le ampie larghezze di banda e le grandi aperture d'antenna disponibili nelle reti mobili di prossima generazione faciliteranno l'uso del radar OFDM per la raccolta di informazioni di rilevamento. 
La combinazione di queste funzioni e delle capacità cellulari sottostanti è denominata Integrated Sensing and Communications (ISAC). 
Le applicazioni tipicamente considerate per l'ISAC includono il rilevamento di veicoli (automobili o droni), il rilevamento di attività umane e la sicurezza industriale. 
In questi casi d'uso il presupposto tipico è che il bersaglio venga illuminato in linea diretta dal segnale radio, condizione denominata \textit{line-of-sight} (\gls{los}). Tuttavia, un ambito non ancora esplorato in modo esaustivo è quello in cui il bersaglio non si trova in linea diretta, ma viene raggiunto dal segnale solo dopo riflessi da altri ostacoli, condizione indicata come \textit{non-line-of-sight} (NLOS).
L'obiettivo di questa tesi è lo studio della fattibilità di sistemi ISAC in condizioni di NLOS e la definizione di un primo approccio per il rilevamento radar in NLOS basato sull'analisi spettrale del periodogramma, affrontando i vincoli sulle prestazioni imposti dai requisiti di comunicazione, dallo sviluppo di algoritmi, dall'identificazione di nuovi scenari applicativi per il sistema.
Gli approcci proposti sono stati convalidati utilizzando misure reali
ottenute dal prototipo ISAC mmWave di Nokia, installato in
una struttura di ricerca industriale ad ARENA 2036, Stoccarda, Germania. 
Il prototipo è basato su un sistema commerciale che opera a 27,4 GHz, conforme alle specifiche del 3rd Generation Partnership Project (3GPP) per le forme d'onda di quinta generazione (5G) per le forme d'onda di quinta generazione (5G) nelle bande di frequenza mmWave. 
\\
\\
\textbf{Parole chiave:} isac, 6g, radar ofdm, radar monostatico, rilevamento nlos % Keywords (italian)

%----------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES/SYMBOLS
%----------------------------------------------------------------------------

% TABLE OF CONTENTS
\thispagestyle{empty}
\tableofcontents % Table of contents 
\thispagestyle{empty}
\cleardoublepage

%-------------------------------------------------------------------------
%	THESIS MAIN TEXT
%-------------------------------------------------------------------------
% In the main text of your thesis you can write the chapters in two different ways:
%
%(1) As presented in this template you can write:
%    \chapter{Title of the chapter}
%    *body of the chapter*
%
%(2) You can write your chapter in a separated .tex file and then include it in the main file with the following command:
%    \chapter{Title of the chapter}
%    \input{chapter_file.tex}
%
% Especially for long thesis, we recommend you the second option.

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\mainmatter % Begin numeric (1,2,3...) page numbering

% --------------------------------------------------------------------------
% NUMBERED CHAPTERS % Regular chapters following
% --------------------------------------------------------------------------

% Template text from model, remove after starting to write
% \include{Template}

% Introduction
\input{Chapters/Introduction}

% Overview of 6G sensing
\input{Chapters/overview}

% Joint comm. and sensing/ theoretical
\input{Chapters/theoretical}

% Processing with TDD frame structure
\input{Chapters/TDDprocessing}

% Radar detection and thresholding techniques
\input{Chapters/radar_detect_threshold}

% First Measurements
\input{Chapters/Tests1}

\input{Chapters/Outlook}

%-------------------------------------------------------------------------
%	BIBLIOGRAPHY
%-------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
% \bibliographystyle{IEEEtran}
\singlespacing
\bibliography{Thesis_bibliography} % The references information are stored in the file named "Thesis_bibliography.bib"
\doublespacing
%-------------------------------------------------------------------------
%	APPENDICES
%-------------------------------------------------------------------------

%\cleardoublepage
\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

% ABBREVIATIONS
\singlespacing
\glsaddall
\printglossaries
\doublespacing


% LIST OF SYMBOLS
% Write out the List of Symbols in this page
\chapter*{List of Symbols} % You have to include a chapter for your list of symbols (


\begin{longtable}[c]{lll}
		\textbf{Variable} & \textbf{Description} & \textbf{SI unit} \\\hline\\[-9px]
		$a$ & complex QAM symbol &  \\[1px]
		$\alpha$ & CFAR constant & \\[1px]
		$b_h$ & magnitude attenuation factor due to target $h$ &  \\[2px]
		$B$ & Bandwidth & Hz \\[1px]
		$c_0$ & speed of light & m/s \\[1px]
		$d$ & distance/range & m \\[2px]
		$d_{res}$ & range resolution & m \\[2px]
		$d_{unamb}$ & unambiguous range & m \\[2px]
		$\eta$ & radar threshold &  \\[2px]
		$\Delta f$ & subcarrier spacing  & Hz \\[2px]	
		$f_C$ & carrier frequency & Hz \\[2px]
		$f_D$ & Doppler shift & Hz \\[2px]
		$\mathbf F$ & CSI matrix &  \\[2px]
		$\mathbf F_{Tx}$, $\mathbf F_{Rx}$ & transmitted/received OFDM frame &  \\[2px]
		$H$ & number of targets illuminated by the signal &  \\[2px]
		$\mu$ & 5G system numerology &  \\[2px]
		$M$ & number of symbols in a OFDM frame &  \\[2px]
		$N$ & number of subcarriers &  \\[2px]
		$\varphi$ & phase term & \\[2px]
		$p_{\text{FA}}$ & probability of false alarm &  \\[2px]
		$p_{\text{FA},bin}$ & probability of false alarm per bin &  \\[2px]
		$Per_{\mathbf F}(n,m)$ & 2D periodogram &  \\[2px]
		$\sigma_n^2$ & estimate of AWGN power &  \\[2px]
		$\tau$ & time delay & s \\[2px]
		$\tau_{max}$ & maximum time delay observable from a target & s \\[2px]
		$\theta$ & azimuth in respect to TX or RX & rad \\[2px]
		$T_{CP}$ & cyclic prefix time & s \\[2px]
		$T_S$ & OFDM symbol time &  s \\[2px]
		$v$ & velocity/speed & m/s \\[2px]
		$v_{res}$ & speed resolution & m/s \\[2px]
		$v_{unamb}$ & unambiguous speed & m/s \\[2px]
		$z$ & time-domain AWGN &  \\[2px]
		$\mathbf W$ & 2D window matrix &  \\[2px]
		$\mathbf Z$ & AWGN matrix &  \\[2px]
\end{longtable}


% LIST OF FIGURES
\listoffigures

% LIST OF TABLES
\listoftables



% ACKNOWLEDGEMENTS
\chapter*{Acknowledgements}
\singlespacing
This thesis work was developed at Nokia Bell Labs Stuttgart, in the Radio Systems Research group.

I would like to thank my thesis supervisor, Professor Maurizio Magarini for introducing me to this opportunity and for his support throughout this project.

I would like to thank the Radio Systems Research team and all the colleagues in Nokia Stuttgart, for their support and assistance in various aspects of this work.
Special thanks to the members of the Sensing Corner$^{\text{TM}}$, especially Dr. Marcus Henninger and Dr. Silvio Mandelli, for their guidance and invaluable support.

I would like to thank my girlfriend Sara for her support and love at all times.

I'm grateful to my dear friend Alessandro St., whom I met upon coming to class on the first day (and hoped to leave university together).

I would like to thank my friends, with whom I shared my university years, Alessandro Sa., Ludovico, Ariele, Flavio, Zeineb, Marta, Riccardo R., Pietro G..

I would like to thank my friends Marcello and Alberto, hoping that we will continue travelling together.

I would also like to thank all my friends I met during my life, Andrea, Edoardo, Giorgio, Stefano, Matteo, Leonardo (Pallo) and Pietro M., Riccardo C..
 
My special gratitude goes to my parents, my sister Anna and my family, for their support during this thesis period, my studies and at all times. I want to dedicate this thesis to them. 


\doublespacing

\cleardoublepage


\end{document}